---
title: "Reproducibility of paper:"
subtitle: "Lin, L., Yi, X., Liu, H. et al. The airway microbiome mediates the interaction between environmental exposure and respiratory health in humans. Nat Med 29, 1750–1759 (2023)."
author: "Javier Mancilla Galindo"
date: "`r Sys.Date()`"
execute: 
  echo: false
  warning: false
toc: true
toc_float: true
format:
  pdf:
    documentclass: scrartcl
editor: visual
---

\pagebreak


#### Packages used

```{r}
#| echo: true
#| message: false
if (!require("pacman", quietly = TRUE)) {
  install.packages("pacman")
}

pacman::p_load(
  tidyverse, # Used for basic data handling and visualization.
  dagitty, #Used in conjunction with https://www.dagitty.net/ to create 
  #directed acyclic graph to inform statistical modelling.  
  lavaan, #Used to create correlation matrix to assess conditional independencies.  
  CBPS, #Used to calculate non-parametric propensity scores for IPW.
  WeightIt, #Used to calculate inverse probability weights.
  boot # Calculate bootstrap confidence intervals. 
)
```

##### Session and package dependencies

```{r}
session <- sessionInfo()
session$BLAS <- NULL
session$LAPACK <- NULL
session$loadedOnly <- NULL

session
```

```{r}
#| include: false

# Create directories for sub-folders 
figfolder <- "../mediation/results/output_figures"
dir.create(figfolder, showWarnings = FALSE)
```

```{r}
#| include: false

# load data 
load("../data/meta_phenotypes.RData")

AMHI <- read.delim("../AMHI calculation Figure 4/AMHI.txt")

Health <- read.delim("../data/meta_airwayHealth_overallHealth.txt")

## Merge dataframes   
meta <- merge(meta.cat, meta.num, by="SampleID")

meta <- merge(meta, Health, by.x = "SampleID", by.y = "X.NAME")

data <- merge(meta, AMHI, by="SampleID")
```

\pagebreak

The summary of criticisms to the paper and rationale for the following analyses are summarized in the [powerpoint presentation](mediation/2024-04-12_Seminar-Mediation-Javier.pptx) contained in the *mediation* directory.

## DAG

The following directed acyclic graph (DAG) was reconstructed based on assumptions available in the paper for the construction of regression models in mediation analyses, by using the [DAGitty website](https://www.dagitty.net/). THe DAG is saved and sourced from the accompanying script ***DAG.r***

```{r}
# load Directed Acyclic Graph ============================================

source("scripts/DAG.r")
plot(DAG)
```

## Testing of conditional independencies in DAG:

This procedure was performed as suggested in [this article](https://doi.org/10.1002/cpz1.45).

Implied conditional independencies:

```{r}
impliedConditionalIndependencies(DAG)
```

```{r}
source("scripts/correlation_matrix.r")
```

```{r}
correlations
```

```{r}
localtests <- localTests(
  DAG, 
  sample.cov = correlations,
  sample.nobs=nrow( subsetcondit ) 
  )

localtests
```

Local tests results plot:

```{r}
#| include: false
png(filename=paste0(figfolder,"/conditional_independencies_original.png"),
    width=8, height=5, units="in", res=300)
plotLocalTestResults(localtests)
dev.off()
```

```{r}
plotLocalTestResults(localtests)
```

I updated the DAG according to the procedure described by [Ankan, 2021](https://currentprotocols.onlinelibrary.wiley.com/doi/full/10.1002/cpz1.45).

```{r}
source("scripts/updated_DAG.r")
plot(DAG_updated)
```

```{r}
localtests <- localTests(
  DAG_updated, 
  sample.cov = correlations,
  sample.nobs=nrow( subsetcondit ) 
  )

localtests
```

Local tests results plot:

```{r}
#| include: false
png(filename=paste0(figfolder,"/conditional_independencies_updated.png"),
    width=8, height=5, units="in", res=300)
plotLocalTestResults(localtests)
dev.off()
```

```{r}
plotLocalTestResults(localtests)
```

```{r}
#| include: false 
rm(DAG,DAG_updated,correlations,localtests)
```

\pagebreak

# Biofuel

```{r}
biofuel <- subsetcondit %>% mutate_at(
    c("Biofuel_exposure",
      "Occupational_pollution",
      "Smoking_binary",
      "SHS_binary",
      "Medication",
      "Gender"),
    factor
  )
```

### Inverse probability weights

Weights for exposure:

```{r}
biofuel$weight1 <- weightit(
  Biofuel_exposure ~ BMI + Medication + Age + Gender + Occupational_pollution,
  biofuel, 
  method = "npcbps",
  over = FALSE)$weights
```

Weights for mediator:

```{r}
biofuel$weight2 <- weightit(
  AMHI ~ BMI + Medication + Age + Gender + Biofuel_exposure +
    Occupational_pollution + Smoking_binary + SHS_binary + year2pm25,
  biofuel, 
  method = "npcbps",
  over = FALSE)$weights
```

Overall weight:

```{r}
biofuel <- biofuel %>% 
  mutate(
  weight = weight1*weight2
  )
```

Confidence interval

```{r}
#| warning: false
source("scripts/confidence_intervals_biofuel.r")
conf_interval
```

```{r}
biofuel_ci_lower <- ifelse(conf_interval$percent[1,4] < 0, 0, conf_interval$percent[1,4])
biofuel_ci_upper <- ifelse(conf_interval$percent[1,5] > 100, 0, conf_interval$percent[1,5])
```

# Occupational pollution

```{r}
occupational <- subsetcondit %>% mutate_at(
    c("Biofuel_exposure",
      "Occupational_pollution",
      "Smoking_binary",
      "SHS_binary",
      "Medication",
      "Gender"),
    factor
  )
```

### Inverse probability weights

Weights for exposure:

```{r}
occupational$weight1 <- weightit(
  Occupational_pollution ~ BMI + Medication + Age + Gender,
  occupational, 
  method = "npcbps",
  over = FALSE)$weights
```

Weights for mediator:

```{r}
occupational$weight2 <- weightit(
  AMHI ~ BMI + Medication + Age + Gender + Biofuel_exposure +
    Occupational_pollution + Smoking_binary + SHS_binary + year2pm25,
  occupational, 
  method = "npcbps",
  over = FALSE)$weights
```

Overall weight:

```{r}
occupational <- occupational %>% 
  mutate(
  weight = weight1*weight2
  )
```

Confidence interval

```{r}
#| warning: false
source("scripts/confidence_intervals_occupational.r")
conf_interval
```

```{r}
occupational_ci_lower <- ifelse(conf_interval$percent[1,4] < 0, 0, conf_interval$percent[1,4])
occupational_ci_upper <- ifelse(conf_interval$percent[1,5] > 100, 0, conf_interval$percent[1,5])
```

# PM 2.5

```{r}
pm25 <- subsetcondit %>% mutate_at(
    c("Biofuel_exposure",
      "Occupational_pollution",
      "Smoking_binary",
      "SHS_binary",
      "Medication",
      "Gender"),
    factor
  )
```

### Inverse probability weights

Weights for exposure:

```{r}
pm25$weight1 <- weightit(
  year2pm25 ~ BMI + Medication + Age + Gender + Smoking_binary + Biofuel_exposure +
    Occupational_pollution,
  pm25, 
  method = "npcbps",
  over = FALSE)$weights
```

Weights for mediator:

```{r}
pm25$weight2 <- weightit(
  AMHI ~ BMI + Medication + Age + Gender + Biofuel_exposure +
    Occupational_pollution + Smoking_binary + SHS_binary + year2pm25,
  pm25, 
  method = "npcbps",
  over = FALSE)$weights
```

Overall weight:

```{r}
pm25 <- pm25 %>% 
  mutate(
  weight = weight1*weight2
  )
```

Confidence interval

```{r}
#| warning: false
source("scripts/confidence_intervals_pm25.r")
conf_interval
```

```{r}
pm25_ci_lower <- ifelse(conf_interval$percent[1,4] < 0, 0, conf_interval$percent[1,4])
pm25_ci_upper <- ifelse(conf_interval$percent[1,5] > 100, 0, conf_interval$percent[1,5])
```

# Smoking

```{r}
smoking <- subsetcondit %>% mutate_at(
    c("Biofuel_exposure",
      "Occupational_pollution",
      "Smoking_binary",
      "SHS_binary",
      "Medication",
      "Gender"),
    factor
  )
```

### Inverse probability weights

Weights for exposure:

```{r}
smoking$weight1 <- weightit(
  Smoking_binary ~ BMI + Medication + Age + Gender + Biofuel_exposure + 
    Occupational_pollution,
  smoking, 
  method = "npcbps",
  over = FALSE)$weights
```

Weights for mediator:

```{r}
smoking$weight2 <- weightit(
  AMHI ~ BMI + Medication + Age + Gender + Biofuel_exposure +
    Occupational_pollution + Smoking_binary + SHS_binary + year2pm25,
  smoking, 
  method = "npcbps",
  over = FALSE)$weights
```

Overall weight:

```{r}
smoking <- smoking %>% 
  mutate(
  weight = weight1*weight2
  )
```

Confidence interval

```{r}
#| warning: false
source("scripts/confidence_intervals_smoking.r")
conf_interval
```

```{r}
smoking_ci_lower <- ifelse(conf_interval$percent[1,4] < 0, 0, conf_interval$percent[1,4])
smoking_ci_upper <- ifelse(conf_interval$percent[1,5] > 100, 0, conf_interval$percent[1,5])
```

# Second hand smoking

```{r}
second_hand_smoking <- subsetcondit %>% mutate_at(
    c("Biofuel_exposure",
      "Occupational_pollution",
      "Smoking_binary",
      "SHS_binary",
      "Medication",
      "Gender"),
    factor
  )
```

### Inverse probability weights

Weights for exposure:

```{r}
second_hand_smoking$weight1 <- weightit(
  SHS_binary ~ BMI + Medication + Age + Gender + Biofuel_exposure + 
    Occupational_pollution + Smoking_binary,
  second_hand_smoking, 
  method = "npcbps",
  over = FALSE)$weights
```

Weights for mediator:

```{r}
second_hand_smoking$weight2 <- weightit(
  AMHI ~ BMI + Medication + Age + Gender + Biofuel_exposure +
    Occupational_pollution + Smoking_binary + SHS_binary + year2pm25,
  second_hand_smoking, 
  method = "npcbps",
  over = FALSE)$weights
```

Overall weight:

```{r}
second_hand_smoking <- second_hand_smoking %>% 
  mutate(
  weight = weight1*weight2
  )
```

Confidence interval

```{r}
#| warning: false
source("scripts/confidence_intervals_second_hand_smoking.r")
conf_interval
```

```{r}
sh_smoking_ci_lower <- ifelse(conf_interval$percent[1,4] < 0, 0, conf_interval$percent[1,4])
sh_smoking_ci_upper <- ifelse(conf_interval$percent[1,5] > 100, 0, conf_interval$percent[1,5])
```

# Table results

```{r}
table_results <- data.frame(
  Exposure = factor(1:5, labels=c(
    "Biofuel_exposure",
    "Occupational_pollution",
    "PM25",
    "Smoking",
    "Secondhand_smoking")
    ),
  ci_lower = c(
    biofuel_ci_lower,
    occupational_ci_lower,
    pm25_ci_lower,
    smoking_ci_lower,
    sh_smoking_ci_lower
  ),
  ci_upper = c(
    biofuel_ci_upper,
    occupational_ci_upper,
    pm25_ci_upper,
    smoking_ci_upper,
    sh_smoking_ci_upper
  ),
  confidence_interval = c(
    paste0(biofuel_ci_lower,", ",round(biofuel_ci_upper,2)),
    paste0(occupational_ci_lower,", ",round(occupational_ci_upper,2)),
    paste0(pm25_ci_lower,", ",round(pm25_ci_upper,2)),
    paste0(smoking_ci_lower,", ",round(smoking_ci_upper,2)),
    paste0(sh_smoking_ci_lower,", ",round(sh_smoking_ci_upper,2))
  )
) %>% mutate(Exposure = fct_rev(Exposure))

table_results 
```

```{r}
ggplot(
  table_results, 
  aes(x=Exposure)
  ) +  
  geom_linerange(
    aes(ymin = ci_lower, ymax = ci_upper),
    colour = "firebrick3",
    size = 2
      ) +
  ylim(c(0,100)) + 
  ylab("Proportion mediated (%)") + 
  coord_flip() +
  theme_bw() + 
  theme(panel.border = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(size=rel(1.2)), 
        axis.text.y = element_text(size=rel(1.2))
  )
```

```{r}
ggsave("Proportion_mediated.png",
       path = figfolder,
       width = 6, height = 4, units = "in", dpi = 300
)
```

\pagebreak

```{r}
#| include: false

# Run this chunk if you wish to clear your environment and unload packages.

pacman::p_unload(negate = TRUE)
```
